### 2024-05-08 slides/introduction.tex
presentation upload

    \label{Minimum Wage}
    \frametitle{State minimum wage trends} % Title
    \framesubtitle{}  % Subtitle
    \rmfamily % Font

    \begin{wideitemize} 
        \item States have been increasing their \textcolor{fblu}{minimum wage rates}, with varying intensity
    \end{wideitemize}

    \begin{center}
        \includegraphics[scale=0.5]{min_wage_plot_simp.png}
    \end{center}

    \hyperlink{min_wage_plot_allstates}{\beamerbutton{All states}}

    \frametitle{Main idea} % Title
    \framesubtitle{}  % Subtitle
    \rmfamily % Font

    \begin{wideitemize}
        \item Opioid misuse can lead the victims to negative labor market outcomes
        \item Minimum wage rates affect these outcomes 
        \vspace{9pt}
        \begin{wideitemize}
            \item[\textcolor{fblu}{\textbullet}] Negatively, through their impact on \textcolor{fblu}{labor demand for low-skill jobs} 
            \item[\textcolor{fblu}{\textbullet}] Positively, if they address the reasons that cause \textcolor{fblu}{addiction} in the first place \(\to\) \textcolor{fblu}{Deaths of despair}
        \end{wideitemize}
        \item The dominance of each effect depends on how \textcolor{fblu}{binding} minimum wage rates are across labor markets
    \end{wideitemize}

@@

    \begin{wideitemize}
        \item Do the effects of \textcolor{fblu}{Opioid Use Disorders} (\textcolor{fblu}{OUDs}) on labor market outcomes interact with minimum wage policies?
        \item If so, how have these policies shaped the ongoing Crisis?
        \item Is the interaction homogeneous across States, or can we expect differences across labor markets?
    \end{wideitemize}

@@

    \frametitle{Policy interactions} % Title
    \framesubtitle{}  % Subtitle
    \rmfamily % Font

    \begin{wideitemize}
        \item More stringent regulations on opioid access can make the victims to 
        \begin{wideitemize}
            \item[\textcolor{fblu}{\textbullet}] \textcolor{fblu}{Quit}, which might imply some disutility, or
            \item[\textcolor{fblu}{\textbullet}] \textcolor{fblu}{Switch to the black market}
        \end{wideitemize}
        \item This decision affects the MPL of the worker
        \item In this scenario, minimum wage increases can
        \vspace{9pt}
        \begin{wideitemize}
            \item[\textcolor{fblu}{\textbullet}] Improve labor outcomes if better living conditions tackle the causes of addiction and relapsing (\(\textcolor{fblu}{w_{min}}\) \textcolor{fblu}{not binding} after the change)
            \item[\textcolor{fblu}{\textbullet}] Foster addiction, if quitting doesn't improve employment prospects because \textcolor{fblu}{minimum wage becomes binding} 
        \end{wideitemize}
    \end{wideitemize}

@@

    \vspace{9pt}
    \textbf{Contribution}: shedding light on how the policy responses to the \textcolor{fblu}{Opioid Crisis} have been affected by \textcolor{fblu}{minimum wage changes}
    \vspace{9pt}

>>>>

    \label{main_idea}
    \frametitle{Main idea} % Title
    \framesubtitle{}  % Subtitle
    \rmfamily % Font

    \begin{wideitemize}
        \item Opioid misuse can lead the victims to negative Labor Market Outcomes (\textcolor{fblu}{LMOs})
        \item The impact of substance abuse on LMOs might be mediated by labor market regulations
        \item \textcolor{fblu}{Minimum wage rates} can affect this impact 
        \vspace{9pt}
        \begin{wideitemize}
            \item[\textcolor{fblu}{\textbullet}] Negatively, through their effect on \textcolor{fblu}{labor demand for low-skill jobs} 
            \item[\textcolor{fblu}{\textbullet}] Positively, if they address the reasons that cause \textcolor{fblu}{addiction} in the first place \(\to\) \textcolor{fblu}{Deaths of despair}
        \end{wideitemize}
        %\item The dominance of each effect depends on how \textcolor{fblu}{binding} minimum wage rates are across labor markets
        \item How \textcolor{fblu}{binding} minimum wage rates are across labor markets is the relevant measure
    \end{wideitemize}

    \hyperlink{min_wage_plot}{\beamerbutton{Minimum wage trends}}
    \hyperlink{min_wage_plot_allstates}{\beamerbutton{All states}}

@@

    \begin{wideitemize}
        \item Do the effects of \textcolor{fblu}{Opioid Use Disorders} (\textcolor{fblu}{OUDs}) on labor market outcomes vary depending on minimum wage bindingness?
        %\item If so, how have these policies shaped the ongoing Crisis?
        %\item Is the interaction homogeneous across States, or can we expect differences across labor markets?
    \end{wideitemize}

@@

    \frametitle{Prescription policies} % Title
    \framesubtitle{}  % Subtitle
    \rmfamily % Font

    \begin{wideitemize}
        \item More stringent regulations on opioid access can make the victims to 
        \vspace{9pt}
        \begin{wideitemize}
            \item[\textcolor{fblu}{\textbullet}] \textcolor{fblu}{Quit}, which might imply some disutility, or
            \item[\textcolor{fblu}{\textbullet}] \textcolor{fblu}{Switch to the black market}
        \end{wideitemize}
        \item This decision affects the MPL of the worker
        \item In this scenario, minimum wage rates might
        \vspace{9pt}
        %\begin{wideitemize}
        %    \item[\textcolor{fblu}{\textbullet}] Improve labor outcomes if better living conditions tackle the causes of addiction and relapsing (\(\textcolor{fblu}{w_{min}}\) \textcolor{fblu}{not binding})
        %    \item[\textcolor{fblu}{\textbullet}] Foster addiction, if quitting doesn't improve labor prospects because \(\textcolor{fblu}{w_{min}}\) \textcolor{fblu}{binding} 
        %\end{wideitemize}
        \begin{wideitemize}
            \item[\textcolor{fblu}{\textbullet}] Increase the rates of unemployment of the workers that don't quit (\(\textcolor{fblu}{w_{min}}\) \textcolor{fblu}{binding})
            \item[\textcolor{fblu}{\textbullet}] Not have an impact on employment rates for addicted workers (\(\textcolor{fblu}{w_{min}}\) \textcolor{fblu}{not binding})
        \end{wideitemize}
    \end{wideitemize}

\begin{comment}
\begin{frame}

    \frametitle{Prescription policies} % Title
    \framesubtitle{}  % Subtitle
    \rmfamily % Font

    \begin{equation*}
        \begin{aligned}
            w^{min}\,\,\text{not binding} & \xrightarrow[\text{passed}]{\text{Prescription law}}
            \begin{cases}
                \text{\textcolor{fblu}{Quit}} \\
                \text{LMOs improve} \\
                \downarrow \text{Prescription rates} \\
                \downarrow \text{Mortality} \\
            \end{cases} \\
            & \\
            w^{min}\,\,\text{binding} & \xrightarrow[\text{passed}]{\text{Prescription law}}
            \begin{cases}
                \text{\textcolor{fblu}{Black market}} \\
                \text{LMOs worsen} \\
                \downarrow \text{Prescription rates} \\
                \uparrow \text{Mortality} \\
            \end{cases} \\ 
        \end{aligned}
    \end{equation*}

\end{frame}
\end{comment}

    \frametitle{Prescription policies} % Title
    \framesubtitle{}  % Subtitle
    \rmfamily % Font

    \begin{wideitemize}
        \item Minimum wage bindingness can affect the decision of the worker to quit opioids
        \item This impact is \textcolor{fblu}{ambiguous} and depends on the \textcolor{fblu}{cost of quitting} and the \textcolor{fblu}{impact of addiction on productivity}
        \item Unless minimum wages force the worker to quit strongly enough, we should see a \textcolor{fblu}{negative impact on employment rates} where the minimum wage is \textcolor{fblu}{more binding}
    \end{wideitemize}

@@

    \vspace{9pt}
    \textbf{Contribution}: shedding light on how the labor effects of the \textcolor{fblu}{Opioid Crisis} have been affected by \textcolor{fblu}{minimum wage rates}
    \vspace{9pt}
